---
title: "Introduction"
author: "Quinn Thomas"
date: "2/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction to Rmarkdown



Be sure to load the following packages

```{r}
library(tidyverse)
library(lubridate)
```

The Reservoir column is different reservoirs in the Roanoke, VA area that a set of water quality variables have been measured.

- Description of data set: <https://portal.edirepository.org/nis/metadataviewer?packageid=edi.200.12>
- Data: <https://pasta.lternet.edu/package/data/eml/edi/200/12/0a62d1946e8d9a511bc1404e69e59b8c>

**Warming-up**

Read the data file into R

```{r}
df <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/12/0a62d1946e8d9a511bc1404e69e59b8c")
```
# Excerise 1

The first exercise is focused on data visualization

This module starts with a visualization exercise so you will start with a filtered dataset named `df2`.  This subsetted dataset is for a particular location (`Site == 50`) for a particular Reservoir (`Reservoir == "FCR"`)

```{r}
df2 <- df |> 
  filter(Reservoir == "FCR" & Site == 50 & Date == as_datetime("2021-05-28 10:48:31"))
```

## Question 1:

- Describe the data set that was imported above (`df`): what are the columns and units? Where do the numbers come from? 
- What is the uncertainty in the measurement? What is the spatial and temporal resolution of the data? What value is used for missing values?  What is the interpretation of missing values?

## Question 2

[Plot](https://r4ds.hadley.nz/data-visualize.html#two-numerical-variables) and *describe* the relationship between temperature and depth.  What might explain the pattern? Hint: use the point geometry

```{r}
#ggplot(df2, aes(x = Depth_m, y = Temp_C)) + geom_point()
```

# Exercise 2

Exercise 2 is focused on cleaning up the full dataset.  Initially, you will be using the full dataset (`df`).  You will be using the `filter`, `select`, `rename`, and `mutate` functions.

## Question 3

We are only interested in the Chla_ugL at the surface of each reservoir. [Filter](https://r4ds.hadley.nz/data-transform.html#filter) the dataset (`df`) to only include measurements less than 1 m deep and [select](https://r4ds.hadley.nz/data-transform.html#sec-select) only the reservoir, site, date, depth, and chlorophyll-a columns


```{r}
df3 <- df |> 
  filter(Depth_m < 1) |> 
  select(c("Reservoir","Site","Date","Depth_m","Chla_ugL"))
```

## Question 4

The column name Date is not an accurate name for the column because it include the time of day as well.  [Rename](https://r4ds.hadley.nz/data-transform.html#rename) the column to `datetime` and save the updated data frame. 

```{r}
#df3 <- df3 |> rename(datetime = Date)
```

## Question 5

In the analysis below, you will need a column that is the year.  Apply the function `year()` from the lubridate package to the `datetime` column and save the updated data frame.

## Question 6

In the analysis below, you will need a column that is the month  Apply the function `month()` from the lubridate package to the `datetime` column and save the updated data frame.

## Question 7

To finalize your cleaned dataset, order the columns by Reservior, Site, Year, and Month.  Hint:  use [`arrange`](https://r4ds.hadley.nz/data-transform.html#arrange)


# Exercise 3

Exercise 3 is focused on high level summary statistics through the use of the `group_by()` and `summarize()` combination.

## Question 8

How many different reservoir are in the dataset and what are their names? Hint: use the [`distinct`](https://r4ds.hadley.nz/data-transform.html#distinct) function

## Question 9

Which reservoir has the highest mean surface Chlorophyll-a? Hint: use `group_by()` and `summarize()` [combination](https://r4ds.hadley.nz/data-transform.html#groups)

```{r}
df |> 
  group_by(Reservoir) |> 
  summarize(mean = mean(Chla_ugL))
```

## Question 10

Which reservoir has the highest observed surface temperature?

# Exercise 4

Exercise 4 dives deeper into the data.   

## Question 11

Plot the surface Chlorophyll-a concentration for each year (averaged across reservoirs) and describe how Chla_ugL is changing over time

## Question 12

Plot the annual mean surface Chlorophyll-a concentration for each year for each reservoir and describe how the patterns differ from the plot in Question 9.   

## Question 13

Plot the mean surface Chlorophyll-a concentration by month for each reservoir and describe how the month of peak Chlorophyll-a is similar or different across the reservoirs.

# Exercise 5

Two of the reservoirs are close in distance to each other (FCR and BVR).  We are interested in how the mean annual surface chlorophyll-a differs between the two and how that difference changes over time.  

## Question 14:

Plot the difference in the annual mean chlorophyll-a between `BVR` and `FCR`.  Hints: you will need to use `group_by`, `summarize`, `pivot_wider`, and `mutate` to get the the cleaned dataset from Exercise 3 into a table that can be used generate this plot.

```{r}
df3 |> 
  mutate(year = year(datetime)) |> 
  filter(Reservoir %in% c("BVR","FCR")) |> 
  group_by(Reservoir, year) |> 
  summarize(Chla_ugL = mean(Chla_ugL, na.rm = TRUE)) |> 
  pivot_wider(names_from = Reservoir, values_from = Chla_ugL) |> 
  mutate(diff = BVR - FCR) |> 
  ggplot(aes(x = year, y = diff)) + geom_line()

```

# Exercise

## Question 15

Saving plot

## Question 16

Write your cleaned data frame to your directory 

## BE SURE TO KNIT YOUR DOCUMENT AND SUBMIT TO CANVAS!

