---
title: "Introduction to Tidyverse Module"
author: "Quinn Thomas"
format:   
  html:
    embed-resources: true
---

This module represent the "Whole Game" for data science in R. In this module you will be importing, tidying, transforming, visualizing, and communicating environmental data.  If you have used Tidyverse previously, then this module is a review.  If you have not used Tidyverse previously then you will use the links to the "R for Data Science" book to learn the skills that you need to complete the module.  

Be sure to load the following packages

```{r}
#| echo: false
library(tidyverse)
library(lubridate)
```

The Reservoir column is different reservoirs in the Roanoke, VA area that a set of water quality variables have been measured.

- Description of data set: <https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=200&revision=13>
- Data: <https://pasta.lternet.edu/package/data/eml/edi/200/13/27ceda6bc7fdec2e7d79a6e4fe16ffdf>

**Warming-up**

Read the data file into R. This file is directly streamed into R memory from a URL, rather than downloaded and saved locally.  Reading from the original sources ensures that you are working with latest and unaltered version

```{r}
file_name <- "https://pasta.lternet.edu/package/data/eml/edi/200/13/27ceda6bc7fdec2e7d79a6e4fe16ffdf"
df <- read_csv(file_name)
```

# Excerise 1

The first exercise is focused on data visualization so you will start with a filtered dataset named `df2`.  This subsetted dataset is for a particular location (`Site == 50`) for a particular Reservoir (`Reservoir == "FCR"`).  In this example `df` is modified and assigned to `df2`

```{r}
df2 <- df |> 
  filter(Reservoir == "FCR" & Site == 50 & DateTime == as_datetime("2022-05-09 12:46:59"))
```

## Question 1:

- Describe the data set that was imported above (`df`) to the best of your ability given the documentation provided: what are the columns and units? Where do the numbers come from? 
- What value is used for missing values?  What is the interpretation of missing values?

## Question 2

[Plot](https://r4ds.hadley.nz/data-visualize.html#two-numerical-variables) and *describe* the relationship between temperature and depth.  What might explain the pattern? Hint: use the point geom

```{r}

```

# Exercise 2

Exercise 2 is focused on cleaning up the full dataset.  Initially, you will be using the full dataset (`df`).  You will be using the `filter`, `select`, `rename`, and `mutate` functions.

## Question 3

We are only interested in the `Chla_ugL` at the surface of each reservoir. [Filter](https://r4ds.hadley.nz/data-transform.html#filter) the dataset (`df`) to only include measurements less than 1 m deep and [select](https://r4ds.hadley.nz/data-transform.html#sec-select) only the reservoir, site, date, depth, chlorophyll-a, chlorophyll-a flag columns.   Remember the filtering is done on rows and selecting is done on columns.  Assign the new dataframe to `df3`

```{r}
df3 <- df |> filter(Depth_m < 1, Reservoir == "FCR") |> select(Reservoir, Site, DateTime, Depth_m, Chla_ugL, Flag_Chla_ugL)
```

## Question 4

The column name `Flag_Chla_ugL` is not is a bit long [Rename](https://r4ds.hadley.nz/data-transform.html#rename) the column to `flag`.  Then `filter` the dataset to only include high quality data (flag is 0) 

```{r}
df3 <- df3 |> rename(flag = Flag_Chla_ugL) |> filter(flag == 0)
```


## Question 5

In the analysis below, you will need a column that is the year of the measurement.  Apply the function [`year()`](https://r4ds.hadley.nz/datetimes.html#date-time-components) from the lubridate package to the `datetime` column. Hint: you will use the `year` function within the [`mutate` function](https://r4ds.hadley.nz/data-transform.html#sec-mutate) 

```{r}

```

## Question 6


In the analysis below, you will need a column that is the month of the measurement.  Apply the function [`month()`](https://r4ds.hadley.nz/datetimes.html#date-time-components) from the lubridate package to the `datetime` column and save the updated data frame. Hint: you will use the [`mutate` function](https://r4ds.hadley.nz/data-transform.html#sec-mutate) 

```{r}

```

## Question 7

To finalize your cleaned dataset, order the values in the columns by Reservoir (A-Z).  Hint:  use [`arrange`](https://r4ds.hadley.nz/data-transform.html#arrange)

```{r}

```


# Exercise 3

Exercise 3 is focused on high level summary statistics.

## Question 8

How many different reservoirs are in the dataset and what are their names? Hint: use the [`distinct`](https://r4ds.hadley.nz/data-transform.html#distinct) function

```{r}

```

## Question 9

Which reservoir has the highest mean surface (< 1 m depth) chlorophyll-a? Hint: use `group_by()` and `summarize()` [combination](https://r4ds.hadley.nz/data-transform.html#groups)

```{r}

```

## Question 10

Which reservoir has the highest observed surface chlorophyll-a?

```{r}

```

# Exercise 4

Exercise 4 dives deeper into the data.   

## Question 11

Plot the surface chlorophyll-a concentration for each year (averaged across reservoirs) and describe how surface chlorophyll-a is changing over time.

```{r}

```

## Question 12

Plot the annual mean surface chlorophyll-a concentration for each year for each reservoir and describe how the patterns differ from the plot in Question 11.  

```{r}

```

## Question 13

Plot the mean surface chlorophyll-a concentration by month for each reservoir and describe how the month of peak chlorophyll-a is similar or different across the reservoirs.

```{r}

```

# Exercise 5

Two of the reservoirs are close in distance to each other (FCR and BVR).  We are interested in how the mean annual surface chlorophyll-a differs between the two and how that difference changes over time. Is the water quality at one reservoir changing more over time than the other? 

## Question 14:

Plot the difference in the annual mean chlorophyll-a between `BVR` and `FCR`.  This is the trickiest of the questions because you will need to use [`pivot_wider`](https://r4ds.hadley.nz/data-tidy.html#widening-data) to make a column for `BVR` and `FCR` so that you can subtract one column from the other.

Hints: you will need to use `group_by`, `summarize`, [`pivot_wider`](https://r4ds.hadley.nz/data-tidy.html#widening-data), and `mutate` to get the cleaned dataset from Exercise 3 into a table that can be used generate this plot.


```{r}

```

# Exercise 6

This exercise is focused on exporting the data and a plot 

## Question 15

Save the plot from Question 14 to a PNG named `question-14.png` in your repo top-level directory. Hint: use [ggsave()](https://r4ds.hadley.nz/data-visualize.html#sec-ggsave)

```{r}

```

## Question 16

Write the data frame you created for answering Question 14 (not the entire dataset) to your repo top-level directory with the filename of `question-14-data.csv`.  Hint use [`write_csv()`](https://r4ds.hadley.nz/data-import.html#sec-writing-to-a-file)

```{r}

```

Remember to Render your document to an html file and comment+push to GitHub your qmd file, rendered html, png (from question 15), and csv (from question 16). 
